Trafficking synaptic materials in complex morphologies involves tradeoffs in speed and precision

### 2015-08-20 latex/paper.tex
finish final 2 figs

\documentclass[fleqn,10pt]{wlpeerj}
\title{Trafficking synaptic materials in complex morphologies involves tradeoffs in speed and precision}

@@

While the position of individual cargoes can be highly stochastic, the net movement of a population of cargoes (Fig. 1C) is predictable. Figure 1D shows the distribution of 1000 molecules over time with (top panel) and without (bottom panel) history-dependence. In both cases, these simulations can be well-fit by a drift-diffusion partial differential equation, which is easier to analyze because it is deterministic and contains fewer variables (\cite{Smith_2001}; Fig. 1E-F).

\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\columnwidth]{00_stochastic.png}
\caption{A simple mass-action model of dendritic transport}
\end{center}
\end{figure}

The partial differential equation in Figure 1E can be approximated by discretizing a dendritic tree into small compartments, and considering the movement of cargo between neighboring compartments as a chemical reaction with first order kinetics. In an unbranched cable with $N$ compartments, the mass-action model is:
\begin{equation}
u_1 \underset{b_1}{\overset{a_1}{\rightleftharpoons}} u_2 \underset{b_2}{\overset{a_2}{\rightleftharpoons}} u_3 \underset{b_3}{\overset{a_3}{\rightleftharpoons}}~...~\underset{b_{N-1}}{\overset{a_{N-1}}{\rightleftharpoons}}  u_N
\end{equation}
where $u_i$ is the mass of cargo in each compartment, and $a_i$ and $b_i$ respectively denote local rate constants for anterograde and retrograde transport. If the compartments have equal volumes, then $u_i$ can equivalently be thought of as the molecule's concentration. Similar mass-action schemes can be constructed to model transport in branched morphologies (Fig. 1A).

A critical assumption of this model is that the movement of molecules between neighboring compartments is a stochastic and memoryless process. This assumption is reasonable if the run length of molecular motors is comparable to the size of the compartments; this appears to be the case under many \citep{Muller_2008,Verbrugge_2009}, thought not necessarily all \citep{Dynes_2007} circumstances (see Discussion). In simulations, we observed that this model was not very sensitive to violations of this assumption; the population dynamics of cargoes with substantial run lengths could still be well-fit by this class of models (Fig. 1F, lower panel; Supp Fig. NUMBER). [SHOULD THIS BE MOVED TO DISCUSSION?]

@@

The exchange of $u$ between neighboring compartments approaches an equilibrium/steady-state ($ss$) distribution over time. In an unbranched cable, this occurs precisely when:
$$
\frac{u_i}{u_{i+1}} \Bigg|_{ss} = \frac{b_i}{a_i}
$$
(see Methods). This result allows us to tune the local trafficking rates, $a_i$ and $b_i$, to produce a desired spatial distribution of molecular cargo. Let $\tilde{u}$ denote the desired steady-state level of $u$ in each compartment, we should select trafficking rate constants that satisfy:
$$
\frac{b_i}{a_i} = \frac{\tilde{u}_i}{\tilde{u}_{i+1}}
$$

For example, we can create an linear gradient by specifying $\tilde{u}_i \propto i$. This results in $b_i / a_i =  i / i+1$. Figure 1B shows simulation data confirming that this model converges to a linear gradient at steady-state. Notice that the slope of the resulting linear gradient depends on the total amount of $u$ present in the cable; the transport rule only re-distributes $u$ so that its steady-state levels are \textit{proportional} to the desired levels (Fig. 1C). If the target distribution is normalized such that $\sum_i \tilde{u}_i = 1$, then the steady-state level in each compartment is given by:
$$
u_i \Big|_{ss} = \tilde{u}_i~ u_{\text{total}}
$$
where $u_{\text{total}} = \sum_i u_i$ is the total amount of $u$ in the neuron. The above results extend directly to branched morphologies (see Methods). For any pair of connected compartments:
$$
\frac{u_p}{u_c} \Bigg|_{ss} = \frac{b}{a}
$$
where $u_p$ is the level in the ``parent" compartment (closer to soma) and $u_c$ is the level in the ``child" compartment (closer to periphery); $b$ and $a$ refer to the trafficking rate constants between this pair of compartments.

The above rules could be locally implemented by simple biochemical reaction networks. Consider the cytoplasmic calcium concentration, $[ca]_i$, as a candidate second messenger. Increases in $[ca]_i$ simultaneously arrest anterograde and retrograde microtubular transport \citep{Wang_2009}. That is, for any pair of compartments, the anterograde rate constant is determined by calcium in the parent compartment, $a = f([ca]_p)$,  and the retrograde rate constant is determined by calcium in the child compartment, $b = f([ca]_c)$, where $f$ is a function that describes how calcium alters the transport rates. This satisfies equation (4) and results in $\tilde{u}_i = 1/f([ca]_i)$:
$$
\frac{b}{a} = \frac{f([ca]_c)}{f([ca]_p)} = \frac{\tilde{u}_p}{\tilde{u}_c}
$$

To test the plausibility of this simple model in a complex morphology, we used an existing reconstruction of a CA1 pyramidal neuron \citep{Migliore_2012} that includes complex voltage-dependent conductances and synapses. Excitatory synaptic input was delivered to 120 locations within three dendritic regions (red dots, Fig. 1D), and the average membrane potential in each electrical compartment determined the target level $\tilde{u}_i$ in each compartment (Methods). This models how molecular cargo could be selectively trafficked to active synaptic sites. Figures 2E and 2F confirm that the spatial distribution of $u$ approaches the desired steady-state.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\columnwidth]{01_mass_action.png}
\caption{Delivery of cargo to active synapses in a complex dendritic tree}
\end{center}
\end{figure}

; the stochastic model in figure 2 is related to this model by a simple rescaling (see Methods), so this analysis is applicable to both cases

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\columnwidth]{03_convergence.png}
\caption{Bottlenecks in transport}
\end{center}
\end{figure}

@@

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\columnwidth]{Fig3.png}
\caption{Multiple strategies for transport in a model including nonuniform microtubular detachment/activation}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.7\columnwidth]{05_heterosynaptic.png}
\caption{Heterosynaptic and non-local interactions.
}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\columnwidth]{tradeoff.png}
\caption{Tradeoff between speed of consolidation and specificity of potentiation.
}
\end{center}
\end{figure}

>>>>

\documentclass[10pt]{wlpeerj}
\usepackage{mathtools}

\title{Trafficking synaptic materials in complex morphologies involves tradeoffs in speed and precision}

@@

While the position of individual cargoes can be highly stochastic, the net movement of a population of cargoes (Fig. 1C) is predictable. Figure 1D shows the distribution of 1000 molecules over time with (top panel) and without (bottom panel) history-dependence. In both cases, these simulations can be well-fit by an convection-diffusion partial differential equation (Fig. 1E-F), which is easier to analyze because it is deterministic and contains fewer variables \citep{Smith_2001}.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{00_stochastic.png}
\caption{A simple mass-action model of dendritic transport}
\end{center}
\end{figure}

The partial differential equation in Figure 1E can be approximated by discretizing a dendritic tree into small compartments, and considering the movement of cargo between neighboring compartments as a chemical reaction with first order kinetics. In an unbranched cable with $N$ compartments, the mass-action model is:
\begin{equation}
u_1 \xrightleftharpoons[b_1]{a_1} u_2 \xrightleftharpoons[b_2]{a_2} u_3 \xrightleftharpoons[b_3]{a_3} ~...~ \xrightleftharpoons[b_{N-1}]{a_{N-1}} u_N
\end{equation}
where $u_i$ is the amount of cargo in each compartment, and $a_i$ and $b_i$ respectively denote local rate constants for anterograde and retrograde transport; these parameters determine the local diffusion and drift coefficients as shown in figure 1E (see Methods). Similar mass-action schemes can be constructed to model transport in branched morphologies (Fig. 2A).

A central assumption of this model is that the movement of molecules between neighboring compartments is a stochastic and memoryless process. This assumption is reasonable if the run length of molecular motors is comparable to the size of the compartments; this appears to be the case under many \citep{Muller_2008,Verbrugge_2009}, thought not necessarily all \citep{Dynes_2007} circumstances (see Discussion). In simulations, we observed that this model was not very sensitive to violations of this assumption; the population dynamics of cargoes with substantial run lengths could still be well-fit by this class of models (Fig. 1F, lower panel; Supp Fig. NUMBER). [SHOULD THIS BE MOVED TO DISCUSSION?]

@@

The exchange of $u$ between neighboring compartments approaches an equilibrium/steady-state ($ss$) distribution over time. In an unbranched cable, this occurs precisely when:
$$
\frac{u_i}{u_{i+1}} \Bigg|_{ss} = \frac{b_i}{a_i}
$$
(see Methods). Thus, the local trafficking rates, $a_i$ and $b_i$, can be tuned to produce a desired spatial distribution of cargo. Let $\tilde{u}$ denote the desired steady-state level of $u$ in each compartment. For example, $\tilde{u}_i$ might refer to the local concentration of biochemical signal that ``tags'' synapses for plasticity (CITE); such a tag might be positively correlated to the desired level of delivered cargo. To match this desired distribution, the ratio of local trafficking rates should satisfy:
$$
\frac{b_i}{a_i} = \frac{\tilde{u}_i}{\tilde{u}_{i+1}}
$$

For example, we can create an linear gradient by specifying $\tilde{u}_i \propto i$ (Fig. 1B-C). The transport rule only determines the steady-state ratios of cargo in neighboring compartments; it cannot control the slope of the linear profile. Instead, the slope is controlled by tuning the total amount of cargo in the cable (Fig. 1C). These results extend directly to branched morphologies (see Methods).

%For any pair of connected compartments:
%$$
%\frac{u_p}{u_c} \Bigg|_{ss} = \frac{b}{a}
%$$
%where $u_p$ is the level in the ``parent" compartment (closer to soma) and $u_c$ is the level in the ``child" compartment (closer to periphery); $b$ and $a$ refer to the local transport rate constants between this pair of compartments.

These rules could be locally implemented by simple biochemical reaction networks. Consider the cytoplasmic calcium concentration, $[ca]_i$, as a candidate second messenger. Increases in $[ca]_i$ simultaneously arrest anterograde and retrograde microtubular transport \citep{Wang_2009}. That is, for any pair of compartments, the anterograde rate constant is determined by calcium in the parent compartment, $a = f([ca]_p)$,  and the retrograde rate constant is determined by calcium in the child compartment, $b = f([ca]_c)$, where $f$ is a function that describes how calcium alters the transport rates. This satisfies equation (4) and results in $\tilde{u}_i = 1/f([ca]_i)$:
$$
\frac{b}{a} = \frac{f([ca]_c)}{f([ca]_p)} = \frac{\tilde{u}_p}{\tilde{u}_c}
$$

To test the plausibility of this model in a complex morphology, we used an existing reconstruction of a CA1 pyramidal neuron \citep{Migliore_2012} that includes complex voltage-dependent conductances and synapses. Excitatory synaptic input was delivered to 120 locations within three dendritic regions (red dots, Fig. 1D), and the average membrane potential in each electrical compartment determined the target level $\tilde{u}_i$ in each compartment (Methods). This models how molecular cargo could be selectively trafficked to active synaptic sites. Figures 2E and 2F confirm that the spatial distribution of $u$ approaches the desired steady-state.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{01_mass_action.png}
\caption{Delivery of cargo to active synapses in a complex dendritic tree}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{03_convergence.png}
\caption{Bottlenecks in transport}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{04_sushi_belt.png}
\caption{Multiple strategies for transport in a model including nonuniform microtubular detachment/activation}
\end{center}
\end{figure}

@@

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{05_heterosynaptic_ca1.png}
\caption{Heterosynaptic and non-local interactions.
}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\columnwidth]{06_heterosynaptic_cable.png}
\caption{Tradeoff between speed of consolidation and specificity of potentiation.}
\end{center}
\end{figure}
